---
title: "Bias Correction using qmap"
author: "Emilio Gozo"
date: "August 9, 2017"
output:
  md_document:
    toc: no
  html_document:
    theme: united
  pdf_document:
    highlight: zenburn
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width=6, fig.height=4, echo = T) 
```

This evaluates the capability of the R package [*qmap*](https://cran.r-project.org/web/packages/qmap/index.html) to do empirical adjustments to the output of a regional climate model simulation. 

### Load the libraries
```{r load_libs}
library(qmap)
library(reshape2)
library(ggplot2) # for plotting
library(cowplot)
```

### Load test data
The observed data:
```{r obs_data}
obs <- read.csv("input/obs_ts.csv")
```
The output of the model simulation:
```{r mod_data}
mod <- read.csv("input/mod_ts.csv")
```

```{r summary_table, echo=FALSE, results='asis'}
a <- data.frame(unclass(summary(obs$val)))
b <- data.frame(unclass(summary(mod$val)))
df_summary <- merge(a, b, by="row.names", all.x=TRUE, sort=F)
names(df_summary) <- c("", "Observed", "Modelled")
knitr::kable(df_summary, caption="**Table 1.** Summary of values")
rm(a, b, df_summary)
```

We see from Table 1 that the modelled result distribution is different from that of the observed.

### Bias Correction using *QmapQuant*
Non-parametric quantile mapping using empirical quantiles can be done using *fitQmapQUANt* and *doQmapQUANT*. *fitQmapQUANT* estimates the [empirical cumulative distribution function](https://en.wikipedia.org/wiki/Empirical_distribution_function) (ECDF) for both the observed and modelled time series for the specified quantiles. This will then be used by *doQmapQUANT* for bias-correction.

Figure 1 illustrates how the mapping is done. To obtain the bias-corected value:  
1. Compute the quantiles of the modelled data set (*modq*)  
2. Compute the quantiles of the observed data set (*fitq*)  
3. Determine the new value by matching the quantile ranks of *modq* with *fitq* 

As an example, the bias-corrected value of 28.7$^\circ$C is 27.07$^\circ$C because they have the same quantile rank.

```{r fitting}
fit <- fitQmapQUANT(obs$val, mod$val)
bc_mod <- doQmapQUANT(mod$val, fit)
```

```{r plot_quant, echo=FALSE, fig.cap=fig_cap, fig.width=7, fig.height=7}
quant <- data.frame(fit$par)
quant$q <- 0:100

gen_segments <- function(ptx, pty, minx, miny) {
    pts <- data.frame(x = ptx, y= pty)
    pts$xend <- pts$x
    pts$yend <- miny
    pts2 <- pts[c('x', 'y')]
    pts2$xend <- minx
    pts2$yend <- pts2$y
    pts <- rbind(pts,pts2)
    return(pts)
}

gen_label <- function(ptx, pty) {
    paste0("(", round(ptx, 2), ", ", round(pty, 2), ")")
}

ptq <- c(25, 87)

gg_line_obs <- ggplot(data=quant, aes(fitq, q))
gg_line_obs <- gg_line_obs + geom_line(col="firebrick", lwd=2)
pts <- gen_segments(unname(quantile(obs$val, ptq/100)), ptq, miny=0, minx=min(obs$val))
gg_line_obs <- gg_line_obs + geom_segment(data=pts, aes(x=x, y=y, xend=xend, yend=yend))
gg_line_obs <- gg_line_obs + geom_point(data=pts, aes(x=x, y=y))
gg_line_obs <- gg_line_obs + geom_text(data=pts, aes(x=x, y=y, label=gen_label(x, y)), vjust=-1.2,  size=3)
gg_line_obs <- gg_line_obs + scale_x_continuous(expand = c(0, 0)) + scale_y_continuous(expand = c(0, 0))
gg_line_obs <- gg_line_obs + xlab(expression("Observed Temperature " ( degree~C)))
gg_line_obs <- gg_line_obs + theme(plot.margin = unit(c(0, 0, 0, 0), "cm"),
                                   text = element_text(size=12))


gg_line_mod <- ggplot(data=quant, aes(q, modq))
gg_line_mod <- gg_line_mod + geom_line(col="tomato", lwd=2)
pts <- gen_segments(ptq, unname(quantile(mod$val, ptq/100)), minx=0, miny=min(mod$val))
gg_line_mod <- gg_line_mod + geom_segment(data=pts, aes(x=x, y=y, xend=xend, yend=yend))
gg_line_mod <- gg_line_mod + geom_point(data=pts, aes(x=x, y=y))
gg_line_mod <- gg_line_mod + geom_text(data=pts, aes(x=x, y=y, label=gen_label(x, y)), vjust=-1.2,  size=3)
gg_line_mod <- gg_line_mod + scale_x_continuous(expand = c(0, 0)) + scale_y_continuous(expand = c(0, 0))
gg_line_mod <- gg_line_mod + ylab(expression("Modelled Temperature " ( degree~C)))
gg_line_mod <- gg_line_mod + theme(axis.text.x = element_text(angle = -90, hjust = 0, vjust = 0.5),
                plot.margin = unit(c(0, 0, 0, 0), "cm"),
                text = element_text(size=12))

gg_scatter <- ggplot(quant, aes(fitq, modq)) + geom_point(col="red", alpha=.8)
pts <- gen_segments(unname(quantile(obs$val, ptq/100)), unname(quantile(mod$val, ptq/100)), minx=min(obs$val), miny=min(mod$val))
gg_scatter <- gg_scatter + geom_segment(data=pts, aes(x=x, y=y, xend=xend, yend=yend))
gg_scatter <- gg_scatter + geom_text(data=pts, aes(x=x, y=y, label=gen_label(x, y)), vjust=-1.2,  size=3)
gg_scatter <- gg_scatter + scale_x_continuous(expand = c(0, 0)) + scale_y_continuous(expand = c(0, 0))
gg_scatter <- gg_scatter + theme(axis.title=element_blank(),
				axis.text=element_blank(),
				plot.margin = unit(c(0, 0, 0, 0), "cm"))

fig_cap <- "Figure 1. QQ plot"
p <- plot_grid(gg_line_mod, gg_scatter, NULL, gg_line_obs, align='hv', labels = c("a", "b", "", "c"), rel_heights=c(5,2), rel_widths=c(2,5))
p
```

Table 2 confirms that the result after bias correction has the same distribution as the observed data set.

```{r summary_table2, echo=FALSE, results='asis'}
a <- data.frame(unclass(summary(obs$val)))
b <- data.frame(unclass(summary(mod$val)))
c <- data.frame(unclass(summary(bc_mod)))
df_summary <- merge(a, b, by="row.names", all.x=TRUE, sort=F)
row.names(df_summary) <- df_summary$Row.names
df_summary$Row.names <- NULL
df_summary <- merge(df_summary, c, by="row.names", all.x=TRUE, sort=F)
names(df_summary) <- c("", "Observed", "Modelled", "BiasCorrected")
knitr::kable(df_summary, caption="**Table 2.** Summary of values")
rm(a, b, c, df_summary)
```

